Rotation Period Evolution in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking

### 2019-07-01 Paper/Response1/response.txt

Neither model reproduces the very supersynchronous, P$_{orb}/$P$_{rot} > 1.6$, binaries in the \citet{Lurie2017} as these binaries have orbital eccentricities in excess of 0.3 and are outside of the region of parameter space we consider. The CTL model, however, could in principle reproduce these points if they are tidally-locked binaries as its P$_{eq}$ is a continuous function of $e$ and is applicable for large $e$, in contrast to the discrete P$_{eq}$ predicted by the CPL model that is valid for smaller $e$.

REF: "4.6 The discussion around spin-orbit resonances as the ones by Rodriguez et al (2012) has already been discussed in much more positive terms in Sect. 2.2.3. + See comment above about the conclusion on CTL or CPL and the comparison to Lurie+2017 data."

REPLY: We removed the repeated discussion of spin-orbit resonances from this section. Furthermore, we cut out additional unnecessary text in this section to make its points more concise.

REF: "4.7 I'm sorry, I don't really understand the point of this section, at best it could be added in the conclusion in a sentence or two about future perspectives."

REPLY: We removed this section and compressed the point to the following few sentences in the discussion as suggested by the reviewer:

Beyond these tests, our model could be used to infer the tidal properties of binary stars, perhaps in a Markov Chain Monte Carlo framework, by directly comparing simulation results with the observed stellar and orbital properties, given the observational uncertainties and reasonable prior probability distributions for parameters like the initial binary $e$. This analysis, however, is also beyond the scope of this work and we leave it for future endeavors.

5. Discussion

REF: "We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries." => Which observational tests ? The discussion would be a good place to summarise them."

REPLY: We added the following text in the discussion to briefly summarize our proposed observational tests:

We described three observational tests that can distinguish between which equilibrium tidal model better describes tidal interactions in low-mass stellar binaries. We primarily suggest that observers measure stellar P$_{rot}$ in binaries with solar-mass primaries at P$_{orb} 60-100$ d. If any tidally-locked binaries are identified at long P$_{orb}$, this would be evidence in favor of the CPL model as we found that only binaries tidally-interacting via the CPL model could tidally-lock at such long P$_{orb}$. At shorter P$_{orb}$, precise measurements of P$_{rot}$ and binary $e$ and P$_{orb}$ could distinguish between the CPL and CTL model in tidally-locked systems, e.g. identifying if P$_{eq}$ follows Eq.~(\ref{eqn:cpl:eqPer}) vs. Eq.~(\ref{eqn:ctl:eqPer}), especially if the CPL P$_{eq}$ is in fact a discrete function of $e$. The observations required by these tests, however, are non-trivial.

Additional comments

REF: "These low-masse stars are expected to have their core and envelope rotation evolution decoupled and to re-couple on a long time-scale, see eg Gallet & Bouvier (2015). How would that affect the results of the authors ? In what direction ? In the aforementioned paper, they seem to have some fast rotating stars at late ages that are non-explained by their single star models. Do the authors think their binary model could help to reproduce these ?"

REPLY: As explained by Gallet et al. (2013) and Gallet et al. (2015), the radiative core could act as an angular momentum reservoir for the envelope for up to 1 Gyr. This effect would work to spin up stars early on the main sequence against the spin-down induced by tidal torques and magnetic braking. We anticipate that this would increase the tidal locking timescales, but not significantly since the transfer lasts for only up to 1 Gyr. One interesting situation that could occur, however, is if the torque from the convective envelope's spin-up from internal angular momentum transportation balances the spin-down from tidal and magnetic braking torques. We would expect that in this case, a binary star could be rotating slightly faster than its expected equilibrium rotation period from tidal locking, analogous to the subsynchronous cases examined in Section 4.1 in this work. This state would likely not persist for long, however, as the internal angular momentum transport does not persist for the age of many systems.

To address this, we added the following text to a new subsubsection, 2.1.1: Core-Envelope Coupling, in the Methods section:

Our simplified model assumes that stars follow solid body rotation, whereas in low-mass stars, coupling between the radiative core and convective envelope can impact the surface rotation period evolution \citep{MacGregor1991,Allain1998,Bouvier2008,Irwin2009}. Recent work by \citet{Gallet2013} and \citet{Gallet2015} find that the rapidly-rotating stellar core acts as an angular momentum reservoir for the convective envelope, potentially transferring angular momentum within the stellar interior and into the envelope for up to 1 Gyr, although this process depends strongly on the adopted magnetic braking model, the initial rotation rate, and the stellar mass. We anticipate that internal angular momentum transport would work against the spin-down caused by tidal torques, increasing tidal locking timescales. Internal angular momentum transport torques could potentially balance both tidal and magnetic braking torques near the tidally-locked state, producing slight supersynchronous rotation, analogous to the subsynchronous case examined in $\S$~\ref{sec:eq}. Modeling core-envelope coupling is beyond the scope of this work, however.

The fast rotating stars at late ages examined by Gallet & Bouvier (2015) could be explained by our model if they are in fact unresolved binaries at relatively short orbital periods. For example, for sufficiently strong tidal parameters, our theory shows that binaries can tidally-lock in good agreement with previous results, e.g. Zahn & Bouchet (1989). In this case, the tidal locking maintains rapid rotation against the spin-down caused by magnetic braking at the expense of orbital angular momentum, often keeping Prot <= 20 d. We find that this state of tidally-induced rapid rotation can persist for Gyrs across a wide range of primary star masses, e.g. Figs. 9, 10, and 14 in the revised manuscript.

Other changes not suggested by the reviewer:

-Throughout the revised manuscript, we corrected typos, grammatical errors, and sentence structure with the corrected text in bold.

- Near the end of the introduction, we performed some minor text edits to improve the concision and removed a redundant phrase. The edited sentence appears in bold in the revised manuscript.

-At the end of Section 2.2.2 in the revised manuscript, we added the following text to further clarify how our results can be interpreted in the context of other similar works as a function of tidal Q:

"For example, a common re-parameterization of Q is the reduced tidal quality factor, $Q' = 3Q/2k_2$ \citep[e.g.][]{Leconte2010}. Given our choice of $k_2 = 0.5$, this reduces to $Q' = 3 Q$.""

-Given that the Matt+2015 model erratum effectively strengthened tides relative to magnetic braking, all the CPL simulations in Fig. 5 in the original manuscript actually tidally-locked after we implemented the correction to the Matt+2015 model instead of rotating subsynchronously, e.g. in the Q = 10^8 case. We re-ran the simulations in Fig. 5 in the revised manuscript, keeping all initial conditions the same, but assuming a slightly longer initial Porb = 7.5 d to get weaker tidal torques. The new simulations now again displays how subsynchronous rotation can occur for short Porb systems and highlights how subsynchronous rotators can transition from subsynchronous to synchronous rotation via the mechanism described in the text. The new simulations qualitatively replicated the evolution seen in original Figure 5 in the submitted manuscript and do not change our results. We noted these changes in bold in the text in Section 4.1.3 and in the Fig. 5 caption in the revised manuscript.

-At the end of the 1st paragraph in Section 4.3 in the revised manuscript, we added the text "consistent with ages of stars in the \kepler field \citep{Chaplin2014}" to clarify our choice of age distribution.

-Following the referee's emphasis on concision, we consolidated much of Section 4.3 to remove redundant points and improve the readability of this section. The scientific content and conclusions remain unchanged, but the language has been edited for clarity. We highlighted the changes in bold text in the revised manuscript.

-------

Again, thank you for the detailed review of our manuscript and we look forward to the next correspondence.

Best,

David Fleming, Rory Barnes, Jim Davenport, and Rodrigo Luger
